Scalar Parametrization for Joint Time-Frequency interpolation in a MIMO-OFDM channel

### 2019-10-11 paper/main.tex
todo changes added

\subsection{Scalar Parametrization}
\label{givens}
The degrees of freedom in the $\bV_{i,t}$ matrices is smaller than the
number of real entries in the matrix because of the geometrical
structure between the columns of the matrix~\cite{4114278}. The
orthonormal matrix $\bV_{i,t}$ with $N_T$ rows and $N_R$ columns can be
decomposed as TODO
\begin{equation}
\textbf{V} = \Bigg[\prod_{k=1}^{n} \bD_{k} \big( \phi_{k,k},\ldots , \phi_{k,n} \big) \: \Bigg\{ \prod_{l=1}^{l=t-k} \bG_{t-l,t-l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{\bI}
\end{equation}
where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k},
\ldots, \phi_{k,t} \big)$ =  $\mbox{diag}\big( {\bf 1}_{k-1},
e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, ${\bf 1}_{k-1}$ is
represents  $k-1$ ones, and the $N_R \times N_T$ matrix $\tilde{\bI}$
= $\big[\bI_{n}, {\bf 0}_{n,t-n}\big]^{T}$, and
\begin{equation}
G_{m-1,m}\big(\theta\big)  =
\begin{bmatrix}
I_{m-2} & & & \\
& \cos\theta & -\sin\theta & \\
& \sin\theta & \cos\theta & \\
& & & I_{t-m}
\end{bmatrix}
\end{equation}
Thus, for a $4 \times 2$ orthogonal matrix $\bV_{i,t}$, we have
\begin{align*}
  \bV_{i,t} & =
  \bD_{1}(\phi_{1,1},\ldots,\phi_{1,4})\bG_{3,4}(\theta_{1,1})
  \bG_{2,3}(\theta_{1,2}) \bG_{1,2}(\theta_{1,3})\times\\
& \bD_{2}(\phi_{2,2},\phi_{2,3},\phi_{2,4}) \bG_{3,4}(\theta_{2,1}) \bG_{2,3}(\theta_{2,2})\tilde{\bI}
\end{align*}
Here, $\phi_{k}$s are referred to as phases and $\theta_{l}$s as
rotation angles. $\phi_k \in (-\pi, + \pi]$ is uniformly distributed
between the two extremes\cite{4114278}, while $\theta_l$
is distributed as
\begin{equation}
p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,t-1
\end{equation}
The $\phi_k$ and $\theta_l$ are statistically independent, making this
representation useful for tracking them separately across time, as
explained in the following subsection. The total number of parameters
obtained from the decomposition of a complex orthogonal matrix
$N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$~\cite{4114278}. The
number of $\phi_i$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the
number of $\theta$ parameters is $N_{T}(2N_{T} - N_{R}+1)/2$.
\begin{figure}
\includegraphics[width=0.45\textwidth]{images/new-adpm.pdf}
\caption{\label{fig:adpm-fig}Subcarriers which need to be quantized}
\label{ber_overvie}
\vspace{-5pt}
\end{figure}
% To exploit the interpolation of time \& frequency jointly at the transmitter, feedback bits are sent in an alternating fashion as shown in the figure.

@@

To quantize the subcarriers efficiently and exploit the joint
time-frequency interpolation at the transmitter side, we select the
subcarriers for quantization in an alternating fashion as shown in
Fig.~\ref{fig:adpm-fig}. The figure shows the channel evolution of an
$N$ subcarrier MIMO-OFDM system with subcarriers which are going to be
quantized at every $s$-th subcarriers, for even time instances, where,
$s$ is the gap between two quantized subcarrier indices. For odd time
instances, subcarriers at the position $sk$ for k = 0,1,...,
$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$. TODO

@@

After obtaining the $\beta$ values(1-bit numbers) at the trasnmitter we will calculate the subcarrier indices at indices sk or sk+t. To reproduce the rest of the subcarriers will use future time instances to do the interpolation. As you can see in the figure, each parameter is interpolated using the subcarriers of the next six time instances. We do the interpolation using the spline-cubic interpolation. We could also use other standard techniques by calculating the MMSE at the missing subcarrier locations but the spline-cubic interpolation method gives better results over the other method.

@@

 We have performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel. Channel is simulated using the typical Jakes Model for both ITU Vehicular-A channel and ITU Pedestrian Channel. Performance is measured by averaging over 10 channel realizations with 100 simulations each. The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with sampling frequency $T_s = 20MHz$. For both vehicular and pedestrian channel, $N_T=4$ and $N_R=2$. Therefore, the total number of parameters are $N_{R}(2N_{T} - N_{R}) = 12$ with 5$\theta$’s and 7$\phi$’s. There are 1024 subcarriers in Vehicular channel and 64 subcarriers in the Pedestrian Channel. We take s = 33/9 for Vehicular and pedestrian channel respectively I.e. 32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1, … , 31 and 0,1, .. ,7 respectively.

@@

{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} . This ratio was decided after trying with different ratios. This implies that if quantied values lose track of the actual values then after changing delta quickly they start following it again.
Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper. We should also try over different quantization rates to compare

While doing interpolation at the transmitter end, future time instances are used to interpolate the rest of the subcarriers indices which are not fed back. To prevent much delay, we are using 6 future time instances and using SciPy cubic interpolation over each parameter in 2D. These interpolated values and then used to reconstruct the precoding matrices all together.

>>>>

\subsection{Scalar Parametrization}
\label{givens}
The degrees of freedom in the $\bV_{i,t}$ matrices is smaller than the
number of real entries in the matrix because of the geometrical
structure between the columns of the matrix~\cite{4114278}. The
orthonormal matrix $\bV_{i,t}$ with $N_T$ rows and $N_R$ columns can be
decomposed as TODO
\begin{equation}
\textbf{V} = \Bigg[\prod_{k=1}^{N_{R}} \bD_{k} \big( \phi_{k,k},\ldots , \phi_{k,N_{R}} \big) \: \Bigg\{ \prod_{l=1}^{l=N_{T}-k} \bG_{N_{T}-l,N_{T} -l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{\bI}
\end{equation}
where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k},
\ldots, \phi_{k,N_T } \big)$ =  $\mbox{diag}\big( {\bf 1}_{k-1},
e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,N_T }}  \big)$, ${\bf 1}_{k-1}$ is
represents  $k-1$ ones, and the $N_R \times N_T$ matrix $\tilde{\bI}$
= $\big[\bI_{N_R }, {\bf 0}_{N_T ,N_T -N_R }\big]^{T}$, and
\begin{equation}
G_{m-1,m}\big(\theta\big)  =
\begin{bmatrix}
I_{m-2} & & & \\
& \cos\theta & -\sin\theta & \\
& \sin\theta & \cos\theta & \\
& & & I_{t-m}
\end{bmatrix}
\end{equation}
Thus, for a $4 \times 2$ orthogonal matrix $\bV_{i,t}$, we have
\begin{align*}
  \bV_{i,t} & =
  \bD_{1}(\phi_{1,1},\ldots,\phi_{1,4})\bG_{3,4}(\theta_{1,1})
  \bG_{2,3}(\theta_{1,2}) \bG_{1,2}(\theta_{1,3})\times\\
& \bD_{2}(\phi_{2,2},\phi_{2,3},\phi_{2,4}) \bG_{3,4}(\theta_{2,1}) \bG_{2,3}(\theta_{2,2})\tilde{\bI}
\end{align*}
Here, $\phi_{k}$s are referred to as phases; their collection is given by $\Phi and $$\theta_{l}$s as rotation angles; their collection is given by $\Theta$. $\phi_k \in (-\pi, + \pi]$ is uniformly distributed
between the two extremes\cite{4114278}, while $\theta_l$
is distributed as
\begin{equation}
p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,N_T -1
\end{equation}
The $\phi_k$ and $\theta_l$ are statistically independent, making this
representation useful for tracking them separately across time, as
explained in the following subsection. The total number of parameters
obtained from the decomposition of a complex orthogonal matrix
$N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$~\cite{4114278}. The
number of $\phi_i$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the
number of $\theta$ parameters is $N_{T}(2N_{T} - N_{R}+1)/2$.
\begin{figure}
\includegraphics[width=0.5\textwidth]{images/new-adpm.pdf}
\caption{\label{fig:adpm-fig}Subcarriers which need to be quantized}
\label{ber_overvie}
\vspace{-5pt}
\end{figure}
% To exploit the interpolation of time \& frequency jointly at the transmitter, feedback bits are sent in an alternating fashion as shown in the figure.

@@

To quantize the subcarriers efficiently and exploit the joint
time-frequency interpolation at the transmitter side, we select the
subcarriers for quantization in an alternating fashion as shown in
Fig.~\ref{fig:adpm-fig}. The figure shows the channel evolution of an
$N$ subcarrier MIMO-OFDM system with subcarriers which are going to be
quantized at every $p$-th subcarriers, for even time instances, where,
$p$ is the gap between two quantized subcarrier indices. For odd time
instances, subcarriers at the position $pk+q$ for k = 0,1,...,
$\frac{N-1}{p}$ with q = ${\frac{p}{2}}$. TODO

@@

After obtaining the $\beta$ values(1-bit numbers) at the trasnmitter we will calculate the subcarrier indices at indices $pk$ or $pk+q$. To reproduce the rest of the subcarriers will use future time instances to do the interpolation. As you can see in the figure, each parameter is interpolated using the subcarriers of the next six time instances. We do the interpolation using the spline-cubic interpolation. We could also use other standard techniques by calculating the MMSE at the missing subcarrier locations but the spline-cubic interpolation method gives better results over the other method.

@@

 We have performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel. Channel is simulated using the typical Jakes Model for both ITU Vehicular-A channel and ITU Pedestrian Channel. Performance is measured by averaging over 10 channel realizations with 100 simulations each. The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with symbol time period $T_d = 5\times10^{-8} s$. For both vehicular and pedestrian channel, $N_T=4$ and $N_R=2$. Therefore, the total number of parameters are $N_{R}(2N_{T} - N_{R}) = 12$ with 5$\theta$’s and 7$\phi$’s. There are 1024 subcarriers in Vehicular channel and 64 subcarriers in the Pedestrian Channel. We take s = 33/9 for Vehicular and pedestrian channel respectively I.e. 32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1, … , 31 and 0,1, .. ,7 respectively.

@@

{ADPM Design}: The decrement/decay rate is given by $\bm$ and increment/growth rate is given by $\bM$ in the Equation~\ref{delta_eqn} . The ratio of $m/M$ is fixed at 2.5 i.e. higher decay than increment of the delta values. This way, quantized values are able to follow the curve more precisely. Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper.

\subsection{Bilinear Interpolation}
Bilinear interpolation is the equivalent of linear interpolation in two dimensions. On the transmitter side, all the scalars - $\Theta$ and $\Phi$ of the missing subcarriers are interpolatted using the values of the neigbours. These interpolated values and then used to reconstruct the precoding matrices all together.
